Prestige bias in cultural evolutionary dynamics

### 2023-03-28 ms/ms.tex
reviewed changes by SE, added comments

% TODO edit so that we focus on indirect success - aka prestige - and decompose it to content based - which we call success - and context based - which we call influence
{Boyd and Richerson} \citep[Ch.~5]{evolutionBook}  suggested that the {assessment} of success can be divided to three categories: \emph{direct bias}, \emph{indirect bias} and \emph{frequency-dependent bias}.
{Direct} bias occurs when one phenotype is more attractive than other phenotypes, and is evaluated by \emph{directly} testing the trait.
For example, an individual observing a ping-pong match can {try} the observed paddle grips to determine which grip is better.
{Frequency}-dependent bias occurs when the probability {of copying} a phenotype is higher or lower than the frequency of the phenotype among demonstrators. 
For example, suppose the common paddle grip is used by 60\% of the demonstrators; if the this grip is adopted by 80\% of copiers, then transmission is under positive frequency bias, also called \emph{conformity}; if it is adopted by 40\% of copiers, then transmission is under negative frequency bias, or \emph{{anti}-conformity} \citep{conformism}. 
The effects of conformity and {anti}-conformity on cultural evolution have been studied with both models \citep{anticonformity,Denton2021,conformity} and experiments \citep{negativeFrequency}.
{Indirect} bias occurs when an individual uses one phenotype to determine the attractiveness of another phenotype.
For example, an observer may copy the paddle grip of the {ping}-pong player who scored more points in the match, thus indirectly evaluating the grip by the points scored.
{However, this} may cause mismatches between the copied trait and the rest of the cultural or genetic repertoire of the individual \citep{Kolodny2022}.
Furthermore, {Boyd and Richerson} \citep[Ch. 8]{evolutionBook} {suggest} that maladaptive traits may spread widely in a population if indirect {bias is} strong enough, e.g., by a runaway process caused by a cultural equivalent of sexual selection \citep{sexualSelectionBook}.
Indeed, helping behaviors can evolve due to horizontal transmission bias even without any benefit to the recipient, or when the benefit is much larger than the cost \citep{cooperation}.

@@

A broader definition of prestige is the ``widespread respect and admiration felt for someone or something on the basis of a perception of their achievements or quality'' (New Oxford American Dictionary).
Indeed, {Chudek et al.}\ \citep{prestige_cultural_learning} have defined prestige bias as ``a tendency to learn from individuals to whom others have preferentially attended, learned or deferred'', and demonstrated its occurrence in in 3-4 year old children.
{Henrich and Gil-White} \citep{prestige_evolution} gave a similar definition of prestige as ``freely conferred deference'', in contrast to \emph{dominance}, and provided examples from the anthropological literature.

To distinguish this form of indirect bias from other definitions of prestige, we call it \emph{influence bias}.
This is an important distinction, as influence is a context-{dependent} bias, rather than a content-{dependent} bias: it does not depend on the phenotype itself but rather on the number of copiers that {have} already copied each role-model, which may be easier and more accurate to estimate {than} success. 
Influence bias also differs from {conformity}, which depend on the frequency of a trait in the population or in a sample of role-models, rather than the social dynamics of copying.

In contemporary human society, social media {make} it especially easy to estimate the social and cultural influence individuals have over others, {which} can have an effect on decision making. Online social networks such as \emph{Facebook} and \emph{Instagram} are known to affect the influence of individuals \citep{social_influence,social_media,influence_analysis}, and specific marketing practices {have been} invented to capitalize on this effect \citep{facebook_marketing}.

In the following, we develop a stochastic model of cultural transmission with prestige bias {that} combines both cross-domain indirect success bias and influence bias.
We develop analytic approximations for this model and analyze its dynamics. We {also} find approximations for the probability and time {to} fixation of a `successful' phenotype (i.e., that is {subject to} success bias). Comparing these approximations to Kimura's approximations for the fixation of a favorable allele \citep{kimura,kimura_average}, we demonstrate that success and influence bias play the role of natural selection and genetic drift, respectively.

@@

We begin with a continuous trait model with indirect bias suggested by {Boyd and Richerson} \citep{evolutionBook}, propose an extension with influence bias, and then develop a model with a dichotomous trait.
We implement our stochastic models and approximations, {perform} statistical analyses, and produce figures using Python \citep{python} with NumPy \citep{numpy} and Matplotlib \citep{mathplotlib}. 
Source code is available at \href{https://github.com/yoavram-lab/PrestigeBias}{https://github.com/yoavram-lab/PrestigeBias}.

\subsection*{Continuous trait}
We follow the {Boyd and Richerson model} \citep{evolutionBook}, assuming only oblique transmission of the trait and omitting the indirect trait in order to reduce model complexity. 
{Consider} a population of $N$ individuals, described by a single trait {that takes continuous values}.
Each generation, $N$ naive individuals, or copiers, choose a single individual from the entire previous generation, or role-models, from which they will copy their trait.
This is different than the usual models studied by Feldman et al \citep{cumul_culture,anticonformity} or Boyd \& Richerson \citep{evolutionBook}, where the population is infinite and each copier samples $n$ role-models, and copies the trait from them. 
Similar to a Wright-Fisher model, {generations are non-overlapping, and}  the entire population is replaced in each generation.
The population at time $t$ can be described by $\vec{A}(t)=\big(A_{1}(t), \ldots, A_{N}(t)\big)$ where $A_{i}(t)$ is trait value of individual $i$ at time $t${, and}  the initial population is drawn from a standard normal distribution, $\vec{A}(0) \sim N(0,1)$ .
Cultural transmission is modeled by a function $F$ such that 
\begin{equation}\label{eq:transmission}
A_{i}(t+1) = F_i(\vec{A}(t)) \;.
\end{equation}

\paragraph{Success bias.}
{Boyd and Richerson} \citep[Ch.~8, p.~247-249]{evolutionBook} describe a  transmission algorithm by defining $F$, a weighted average of the traits of all role-models, as
\begin{equation}\label{eq:boydF}
F_i(\vec{A}) = \sum_{j=1}^N G_{i,j}\cdot A_{i,j} \;, 
\end{equation}
where $G_{i,j}$ is the success bias of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:boydG}
G_{i,j} = \frac{\beta(A_{i,j})}{\sum_{k=1}^{N} \beta(A_{i,k})} \;,
\end{equation}
$A_{i,j}$ is the absolute trait value {that} copier $i$ estimates for role-model $j$ with some error $e_i \sim N(0,\eta^2)$,
\begin{equation}\label{eq:relativeTrait}
A_{i,j} = A_j + e_i,
\end{equation}
and $\beta(\cdot)$ is the bias function that quantifies the success bias of a role-model \citep[eq.~5.11]{evolutionBook},
\begin{equation}\label{eq:success_bias}
\beta(A_{i,j}) = b \cdot \exp{\Big(-\frac{(A_{i,j} - \hat{A})^2}{2J}\Big)} \;,
\end{equation} 
with $\hat{A}$ as the arbitrary optimal trait value, and $J$ and $b$ as parameters that control the bias strength.
Therefore, $G_{i,j}$ is a relative success score that copier $i$ assigns to role-model $j$.

{Boyd and Richerson} \citep{evolutionBook} note that the deterministic blended transmission algorithm they use has alternatives. We {can} develop a similar stochastic model with transmission from a single random role-model{\ where instead} of \cref{eq:boydF}, we define the transmission function $F$ as a random variable with its distribution given by 
\begin{equation}\label{eq:ourF}
\Pr\big(F_i(\vec{A}) = A_{j}\big) = G_{i,j} \;;
\end{equation}
{here $G_{i,j}$ is} the probability {that copier $i$ chooses} to copy the trait of role-model $j$.

\paragraph{Influence bias.}
Here we introduce a new element to the model {by assuming} that in each generation, copiers choose their role-models one by one.
{Denote by} $K_{i,j}$  the number of copiers that {choose} role-model $j$ after copier $i$ chose a role-model. Thus, $i$ out of $N$ copiers {had} already {chosen} a role-model, $\sum_{j=1}^N{K_{i,j}} = i$, and there are $N-i$ copiers that have yet to choose a role-model.
The stochastic process of role-model choice, 
\begin{equation} \label{eq:process}
\big\{\vec{K}_i = (K_{i,1}, \ldots, K_{i,N}) \big\}_{i=1}^N \;,
\end{equation}
is described by the recurrence equation
\begin{equation} \label{eq:recurrence}
K_{i,j} = K_{i-1,j} + S_{i,j}, \quad i,j=1,2,\ldots,N \;,
\end{equation}
where $S_{i,j}=1$ if the $i$-th copier chose role-model~$j$ and 0 otherwise, and the initial state is $K_{0,j}=0$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
Our models are defined by two nested stochastic processes. Change over multiple generations is described by the {dynamics of the} phenotype distribution at each generation, $\{\vec{A}(t)\}_t$, see \cref{eq:transmission}. The transition from one generation to the next is described by the number of copiers each role-model has after $i$ copiers have chosen a role-model, $\{\vec{K}_i\}_{i=1}^N$, see \cref{eq:process}.
We emphasize that the models are nested: $\vec{A}(t+1)$ can be computed from $\vec{A}(t)$ by evaluating $\vec{K}_{N}$, where $K_{N,j}$ is the number of copiers that chose role-model $j$ after all copiers chose a role model.
However, the former requires iterating over \cref{eq:recurrence,eq:recPrestige}.
Thus, we sought to find an equivalent stochastic process that has the same joint distribution as $\vec{K}_{N}$. 
We found two approximations for the distribution of $\vec{K}_{N}$, summarized here and explained in detail below.
In both we assume that the bias weight is either completely homogeneous, $\alpha_{i,j}=\alpha$, or that $\alpha_j$ is a bias of role-model $j$ only, meaning it does not vary between copiers.
Note that these approximations apply for both the and the continuous trait (\cref{eq:prestige}) and the dichotomous trait (\cref{eq:binary-model}) models.

@@

\begin{proof}
{Let} $Q_j(k,i)=P(K_{i,j} = k \mid K_{i-1,j})$ {be} the probability that exactly $k$ out of $i$ copiers choose role-model $j$ given $K_{i-1,j}$ {out} of $i-1$ copiers chose role-model $j$. 
Using conditional probability and \cref{eq:recurrence},
\begin{equation}\label{recursive}
Q_j(k,i) = P_j(S_{i,j}=1 | k-1,i-1) \cdot Q_j(k-1,i-1) + P_j(S_{i,j} =0 | k,i-1) \cdot Q_j(k,i-1) \;,
\end{equation}
where $S_{i,j} =1 $ when the $i$-th copier chooses role-model $j$.
\Cref{recursive} is equivalent to eq.~(2.1) by \citep{GBD}, which completes the proof.
\end{proof}
This result gives the following two corollaries on the expected number of followers of a given role-model $j$ by the end of the role-model choice process, $K_{N,j}$. The proof of \Cref{cor:GBD} is in Appendix~\ref{sec:GBD}.
\\

@@

\begin{proof}
{Drezner and Farnum} \citep[eq.~(2.3)]{GBD} show that the expected number of successful trials, 
$\E[k] = N \cdot Q_j(1,1)$ (modified from their notation).
$Q_j(1,1)$ is the initial probability to choose role-model $j$, before any role-model choices are made.
$Q_j(1,1) = G_{1,j}$ by definition, which completes the proof.
\end{proof}

The special case where the bias weight is uniform for all role-models ($\alpha = \alpha_j$ for $j=1,\ldots, N$) is interesting, because we can evaluate the expected number of copiers using a linear equation
\begin{equation}
\E[K_{Nj}]= N\cdot\frac{\alpha\cdot\beta(A_j+e)}{\sum\limits_{m=1}^{N}\alpha\cdot\beta(A_m+e)} =\beta(A_j+e) \bigg/ \overline{\beta(A+e)} \;,
\end{equation}
where the only variable is $A_j+e$, because $\overline{\beta(A+e)}$ is the mean of the distribution of the trait values, modified by some constant parameters of $\beta$.
We can then {write} $L = 1/\overline{\beta(A+e)}$ and
\begin{equation}\label{eq:linearEq}
\E[K_{Nj}] = L\cdot \beta(A_j+e) \;.
\end{equation}

\paragraph{Numerical validation.}
To validate that our approximation is correct
we ran $1,000$ simulations of the full model, and compared the results with \Cref{cor:GBD}.
We compare the distribution of number of copiers by plotting the histograms of both our simulations results and the expected values based on \Cref{cor:GBD}.

@@

Frigyik et al.~\citep[section 2]{dirichlet} prove that the proportion of different colored balls in a \emph{P\'{o}lya urn model} converges to the Dirichlet distribution as the number of draws approaches infinity, $N$ in our model,
based on the \emph{Martingale Convergence Theorem} \citep{martingaleBook}.
We therefore have an approximation for the relative prestige each role-model has when evaluated by copiers. Thus, choosing the role-models for all copiers is equivalent to drawing from a {multinomial} distribution where the parameters are the modified weights from a Dirichlet distribution and we have the following corollary.
\\

@@

\paragraph{Numerical validation.}
To validate our model can be approximated by the Dirichlet distribution properly, i.e  (\Cref{cor:dirichlet}),
and test its sensitivity to the assumptions ($e_i=e$ and $\alpha_i=\alpha$ for $i=1,\ldots,N$) we compare it to results of stochastic simulations of the model, meaning we compare \Cref{cor:dirichlet} to  \Cref{eq:binary-model}.
First, we compute an observed distribution of the number of copiers from the average empirical distribution of multiple simulations.
We then {compare} this observed distribution with the expected theoretical DM distribution (\Cref{fig:DM_validation}A). The difference in distributions was not perceived when plotting both distributions on the same figure, so we used the difference instead. The maximum difference is 0.5 role-models, which {indicates} a very good fit.
In addition, we tested the likelihood of the observed data to be drawn from the DM distribution, against a shuffle of the parameters vector of the DM distribution itself (\Cref{fig:DM_validation}B). We see that the negative log likelihood of the observed data is much higher than any other shuffled version of the parameters vector, {strongly} supporting our approximation.

Next, we examined the fixation probability and fixation time of a favored phenotype $\hat{A}$ when invading a population of phenotype $A$ and compared results from the full model and the DM approximation.
Thus, we assume the population has a single individual with phenotype $\hat{A}$ and $N-1$ individuals with phenotype $A$. 
That the number of simulations needed to sufficiently approximate our model with the DM approximation is roughly $1,000$ (\Cref{fig:num_sims}).
We examined the robustness of the DM approximation to relaxing the following assumptions.
First, we relaxed our assumption of no estimation error $e$.
Estimation error in the original model was drawn from a normal distribution, and added to the trait value before evaluation of the bias ($A_{ij} = A_j + e_i$).
When estimation error is applied, we sample $J_i$ for each copier $i$ from a normal distribution.
Even when the standard deviation is $0.1$, the fixation probability and time {are} similar~(\Cref{fig:hetro_error}). 
We also relaxed our assumption of a uniform bias weight $\alpha$ (i.e., $\alpha_i=\alpha$). We allowed $\alpha$ to vary in the population, drawing $\alpha_j$ for each role-model $j$ from a normal distribution such that $\alpha_j \sim N(0.5,q)$ where $q \in [10^{-7},10^{-1}]$. 
We found again that results of the DM approximation are similar to those from stochastic simulations of the full model~(\Cref{fig:hetro_alpha}).

@@

\begin{result}[Drift and diffusion terms in a constant environment]
Let $x$ and $x'$ be the frequency of type $\hat{A}$ in a population with $N$ individuals in the current and next generation, and  $\beta$ {be} the success coefficient of phenotype $A$, $\beta = \beta(A) < \beta(\hat{A}) = 1$.

Then,
\begin{equation} 
E[x'-x] = \mathcal{O} (x(1-x)(1-\beta)) \;, 
\quad
V(x'-x) = \mathcal{O} (x(1-x)\Big(\frac{1}{\alpha N + (1-\alpha)}\Big)) \;.
\end{equation} 
\end{result}

@@

\begin{result}[Fixation time] The {expected} fixation time (conditioned on fixation) from an initial frequency $x$ is approximately
\begin{equation} \label{eq:kimura_t}
T(x) = J_1(x) + \frac{1-\pi(x)}{\pi(x)}\cdot J_2(x),
\end{equation}
where $N_e=\alpha N + (1-\alpha)$, $S=N_e(1-\beta)$, and
\begin{equation}
\begin{aligned}
J_1(x) &= \frac{-1}{(1-\beta)(e^{-2S}-1)}\int_x^1 \frac{1-e^{2S\xi}-e^{-2S(1-\xi)}+e^{-2S}}{\xi(1-\xi)}d\xi \;, \\
J_2(x) &= \frac{-1}{(1-\beta)(e^{-2S}-1)}\int_0^x \frac{(1-e^{2S\xi})(e^{-2S\xi}-1)}{\xi(1-\xi)}d\xi \;.
\end{aligned}
\end{equation}
\end{result}
Note that these integrals cannot be solved in closed form, {and are estimated} numerically.

@@

\paragraph{Numerical validation.}
To validate that \Cref{eq:ch_env} can approximate our dichotomous model, we compared the results of this equation against simulations using \Cref{cor:dirichlet}. We find that the approximation fits the simulation results well for variable bias weights, $\alpha$, which corresponds to the effective population size (\Cref{fig:ch_env_alpha_beta}A).

>>>>

{Boyd and Richerson} \citep[Ch.~5]{evolutionBook} suggested that the {assessment} of success can be divided to three categories: \emph{direct bias}, \emph{indirect bias} and \emph{frequency-dependent bias}.
{Direct} bias occurs when one phenotype is more attractive than other phenotypes, and is evaluated by \emph{directly} testing the trait.
For example, an individual observing a ping-pong match can {try} the observed paddle grips to determine which grip is better.
{Frequency}-dependent bias occurs when the probability {of copying} a phenotype is higher or lower than the frequency of the phenotype among demonstrators. 
For example, suppose the common paddle grip is used by 60\% of the demonstrators; if the this grip is adopted by 80\% of copiers, then transmission is under positive frequency bias, also called \emph{conformity}; if it is adopted by 40\% of copiers, then transmission is under negative frequency bias, or \emph{{anti}-conformity} \citep{conformism}. 
The effects of conformity and {anti}-conformity on cultural evolution have been studied with both models \citep{anticonformity,Denton2021,conformity} and experiments \citep{negativeFrequency}.
{Indirect} bias occurs when an individual uses one phenotype to determine the attractiveness of another phenotype.
For example, an observer may copy the paddle grip of the {ping}-pong player who scored more points in the match, thus indirectly evaluating the grip by the points scored.
{However, this} may cause mismatches between the copied trait and the rest of the cultural or genetic repertoire of the individual \citep{Kolodny2022}.
Furthermore, {Boyd and Richerson} \citep[Ch. 8]{evolutionBook} {suggest} that maladaptive traits may spread widely in a population if indirect {bias is} strong enough, e.g., by a runaway process caused by a cultural equivalent of sexual selection \citep{sexualSelectionBook}.
Indeed, helping behaviors can evolve due to horizontal transmission bias even without any benefit to the recipient, or when the benefit is much larger than the cost \citep{cooperation}.

@@

A broader definition of prestige is the ``widespread respect and admiration felt for someone or something on the basis of a perception of their achievements or quality'' (New Oxford American Dictionary).
Indeed, Chudek et al.~\citep{prestige_cultural_learning} have defined prestige bias as ``a tendency to learn from individuals to whom others have preferentially attended, learned or deferred'', and demonstrated its occurrence in in 3-4 year old children.
{Henrich and Gil-White} \citep{prestige_evolution} gave a similar definition of prestige as ``freely conferred deference'', in contrast to \emph{dominance}, and provided examples from the anthropological literature.

To distinguish this form of indirect bias from other definitions of prestige, we call it \emph{influence bias}.
This is an important distinction, as influence is a context-dependent bias, rather than a content-{dependent} bias: it does not depend on the phenotype itself but rather on the number of copiers that have already copied each role-model, which may be easier and more accurate to estimate than success. 
%TODO check this - is influence frequency dependent?
Influence bias also differs from conformity, which depend on the frequency of a trait in the population or in a sample of role-models, rather than the social dynamics of copying.

In contemporary human society, social media make it especially easy to estimate the social and cultural influence individuals have over others, which can have an effect on decision making. Online social networks such as \emph{Facebook} and \emph{Instagram} are known to affect the influence of individuals \citep{social_influence,social_media,influence_analysis}, and specific marketing practices have been invented to capitalize on this effect \citep{facebook_marketing}.

In the following, we develop a stochastic model of cultural transmission with prestige bias that combines both cross-domain indirect success bias and influence bias.
We develop analytic approximations for this model and analyze its dynamics. We also find approximations for the probability and time to fixation of a `successful' phenotype (i.e., that is {subject to} success bias). Comparing these approximations to Kimura's approximations for the fixation of a favorable allele \citep{kimura,kimura_average}, we demonstrate that success and influence bias play the role of natural selection and genetic drift, respectively.

@@

We begin with a continuous trait model with indirect bias suggested by Boyd and Richerson~\citep{evolutionBook}, propose an extension with influence bias, and then develop a model with a dichotomous trait.
We implement our stochastic models and approximations, perform statistical analyses, and produce figures using Python \citep{python} with NumPy \citep{numpy} and Matplotlib \citep{mathplotlib}. 
Source code is available at \href{https://github.com/yoavram-lab/PrestigeBias}{https://github.com/yoavram-lab/PrestigeBias}.

\subsection*{Continuous trait}
%TODO If you omit the indirect trait, it is no longer indirect transmission.
We follow the Boyd and Richerson model~\citep{evolutionBook}, assuming only oblique transmission of the trait and omitting the indirect trait in order to reduce model complexity. 
Consider a population of $N$ individuals, described by a single trait that takes continuous values.
Each generation, $N$ naive individuals, or copiers, each chooses a single role-model from the entire previous generation. The copier than copies its trait value from the chosen role-model.
Note that our transmission models is slightly different than previous those modeled before, e.g. \citep{cumul_culture,anticonformity,evolutionBook}, in which the population is infinite and each copier samples $n$ role-models and then copies its trait from one or more of the sampled role-models.

Similar to a Wright-Fisher model, generations are non-overlapping, and the entire population is replaced in each generation.
The population at time $t$ can be described by $\vec{A}(t)=\big(A_{1}(t), \ldots, A_{N}(t)\big)$ where $A_{i}(t)$ is trait value of individual $i$ at time $t$, and  the initial population is drawn from a standard normal distribution, $\vec{A}(0) \sim N(0,1)$ .
Cultural transmission is modeled by a function $F$ such that 
\begin{equation}\label{eq:transmission}
A_{i}(t+1) = F_i(\vec{A}(t)) \;.
\end{equation}

\paragraph{Success bias.}
Boyd and Richerson~\citep[Ch.~8, p.~247-249]{evolutionBook} describe a  transmission algorithm by defining $F$, a weighted average of the traits of all role-models, as
\begin{equation}\label{eq:boydF}
F_i(\vec{A}) = \sum_{j=1}^N G_{i,j}\cdot A_{i,j} \;, 
\end{equation}
where $G_{i,j}$ is the success bias of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:boydG}
G_{i,j} = \frac{\beta(A_{i,j})}{\sum_{k=1}^{N} \beta(A_{i,k})} \;,
\end{equation}
$A_{i,j}$ is the absolute trait value {that} copier $i$ estimates for role-model $j$ with some error $e_i \sim N(0,\eta^2)$,
\begin{equation}\label{eq:relativeTrait}
A_{i,j} = A_j + e_i,
\end{equation}
and $\beta(\cdot)$ is the bias function that quantifies the success bias of a role-model \citep[eq.~5.11]{evolutionBook},
\begin{equation}\label{eq:success_bias}
\beta(A_{i,j}) = b \cdot \exp{\Big(-\frac{(A_{i,j} - \hat{A})^2}{2J}\Big)} \;,
\end{equation} 
with $\hat{A}$ as the arbitrary optimal trait value, and $J$ and $b$ as parameters that control the bias strength.
Therefore, $G_{i,j}$ is a relative success score that copier $i$ assigns to role-model $j$.

Boyd and Richerson~\citep{evolutionBook} note that the deterministic blended transmission algorithm they use has alternatives. We {can} develop a similar stochastic model with transmission from a single random role-model where instead of \cref{eq:boydF}, we define the transmission function $F$ as a random variable with its distribution given by 
\begin{equation}\label{eq:ourF}
\Pr\big(F_i(\vec{A}) = A_{j}\big) = G_{i,j} \;;
\end{equation}
{here $G_{i,j}$ is} the probability {that copier $i$ chooses} to copy the trait of role-model $j$.

\paragraph{Influence bias.}
Here we introduce a new element to the model by assuming that in each generation, copiers choose their role-models one by one.
{Denote by} $K_{i,j}$  the number of copiers that {choose} role-model $j$ after copier $i$ chose a role-model. Thus, $i$ out of $N$ copiers {had} already {chosen} a role-model, $\sum_{j=1}^N{K_{i,j}} = i$, and there are $N-i$ copiers that have yet to choose a role-model.
The stochastic process of role-model choice, 
\begin{equation} \label{eq:process}
\big\{\vec{K}_i = (K_{i,1}, \ldots, K_{i,N}) \big\}_{i=1}^N \;,
\end{equation}
is described by the recurrence equation
\begin{equation} \label{eq:recurrence}
K_{i,j} = K_{i-1,j} + S_{i,j}, \quad i,j=1,2,\ldots,N \;,
\end{equation}
where $S_{i,j}=1$ if the $i$-th copier chose role-model~$j$ and 0 otherwise, and the initial state is $K_{0,j}=0$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
Our models are defined by two nested stochastic processes. Change over multiple generations is described by the {dynamics of the} phenotype distribution at each generation, $\{\vec{A}(t)\}_t$, see \cref{eq:transmission}. The transition from one generation to the next is described by the number of copiers each role-model has after $i$ copiers have chosen a role-model, $\{\vec{K}_i\}_{i=1}^N$, see \cref{eq:process}.
We emphasize that the models are nested: $\vec{A}(t+1)$ can be computed from $\vec{A}(t)$ by evaluating $\vec{K}_{N}$, where $K_{N,j}$ is the number of copiers that chose role-model $j$ after all copiers chose a role model.
However, the former requires iterating over \cref{eq:recurrence,eq:recPrestige}.
Thus, we sought to find an equivalent stochastic process that has the same joint distribution as $\vec{K}_{N}$. 
We found two approximations for the distribution of $\vec{K}_{N}$, summarized here and explained in detail below.
In both we assume that the bias weight is either completely homogeneous, $\alpha_{i,j}=\alpha$, or that $\alpha_j$ is a bias of role-model $j$ only, meaning it does not vary between copiers.% TODO but result 1 has alpha_i
Note that these approximations apply for both the and the continuous trait (\cref{eq:prestige}) and the dichotomous trait (\cref{eq:binary-model}) models.

@@

\begin{proof}
{Let} $Q_j(k,i)=P(K_{i,j} = k \mid K_{i-1,j})$ {be} the probability that exactly $k$ out of $i$ copiers choose role-model $j$ given $K_{i-1,j}$ {out} of $i-1$ copiers chose role-model $j$. 
Using conditional probability and \cref{eq:recurrence},
\begin{equation}\label{recursive}
Q_j(k,i) = P_j(S_{i,j}=1 | k-1,i-1) \cdot Q_j(k-1,i-1) + P_j(S_{i,j} =0 | k,i-1) \cdot Q_j(k,i-1) \;,
\end{equation}
where $S_{i,j} =1 $ when the $i$-th copier chooses role-model $j$.
\Cref{recursive} is equivalent to eq.~(2.1) in \citep{GBD}, which completes the proof.
\end{proof}

This result gives the following corollary on the expected number of followers of a given role-model $j$ by the end of the role-model choice process, $K_{N,j}$.
The proof of \Cref{cor:GBD} is in Appendix~\ref{sec:GBD}.
\\

@@

\begin{proof}
Drezner and Farnum~\citep[eq.~(2.3)]{GBD} show that the expected number of successful trials, 
$\E[k] = N \cdot Q_j(1,1)$ (modified from their notation). % TODO what is k? not defined.
$Q_j(1,1)$ is the initial probability to choose role-model $j$, before any role-model choices are made.
$Q_j(1,1) = G_{1,j}$ by definition, which completes the proof.
% TODO where is the proof for the second part of the corollary??
\end{proof}

The special case where the bias weight is uniform for all role-models ($\alpha = \alpha_j$ for $j=1,\ldots, N$) is interesting, because we can evaluate the expected number of copiers using a linear equation
\begin{equation}
\E[K_{Nj}]= N\cdot\frac{\alpha\cdot\beta(A_j+e)}{\sum\limits_{m=1}^{N}\alpha\cdot\beta(A_m+e)} =\beta(A_j+e) \bigg/ \overline{\beta(A+e)} \;,
\end{equation}
where the only variable is $A_j+e$, because $\overline{\beta(A+e)}$ is the mean of the distribution of the trait values, modified by some constant parameters of $\beta$.
We can then {write} $L = 1/\overline{\beta(A+e)}$ and
\begin{equation}\label{eq:linearEq}
\E[K_{N,j}] = L\cdot \beta(A_j+e) \;.
\end{equation}

\paragraph{Numerical validation.}
To validate that our approximation is correct, % TODO which approximation?? be explicit. refer to an equation. tell us what you are approximating.
we ran $1,000$ simulations of the full model, and compared the results with \Cref{cor:GBD}.
We compare the distribution of number of copiers by plotting the histograms of both our simulations results and the expected values based on \Cref{cor:GBD}.

@@

Frigyik et al.~\citep[section 2]{dirichlet} prove that the proportion of different colored balls in a \emph{P\'{o}lya urn model} converges to the Dirichlet distribution as the number of draws approaches infinity (the population size, $N$, in our model), % TODO did we check sensitivity of the approximation to small values of N? if so, refer to it here.
based on the \emph{Martingale Convergence Theorem} \citep{martingaleBook}.
We therefore have an approximation for the relative prestige each role-model has when evaluated by copiers. Thus, choosing the role-models for all copiers is equivalent to drawing from a {multinomial} distribution where the parameters are the modified weights from a Dirichlet distribution and we have the following corollary.
\\

@@

\paragraph{Numerical validation.}
We next validated the DM approximation of our model
and tested its sensitivity to the assumptions ($e_i=e$ and $\alpha_i=\alpha$ for $i=1,\ldots,N$) by comparing results of stochastic simulations of our model (\cref{eq:binary-model}) with the DM approximation (\Cref{cor:dirichlet}).
First, we computed an observed distribution of the number of copiers from the average empirical distribution of multiple simulations.
We then compared this observed distribution with the expected theoretical DM distribution (\Cref{fig:DM_validation}A).
The difference in distributions was not perceived when plotting both distributions on the same figure, so we used the difference instead.
The maximum difference is 0.5 role-models, which indicates a very good fit.
In addition, we tested the likelihood of the observed data to be drawn from the DM distribution, against a shuffle of the parameters vector of the DM distribution itself (\Cref{fig:DM_validation}B). We see that the negative log likelihood of the observed data is much higher than any other shuffled version of the parameters vector, strongly supporting our approximation.

Next, we examined the fixation probability and fixation time of a favored phenotype $\hat{A}$ when invading a population of phenotype $A$ and compared results from the full model and the DM approximation.
Thus, we assume the population has a single individual with phenotype $\hat{A}$ and $N-1$ individuals with phenotype $A$. 
That the number of simulations needed to sufficiently approximate our model with the DM approximation is roughly $1,000$ (\Cref{fig:num_sims}).
We examined the robustness of the DM approximation to relaxing the following assumptions.
First, we relaxed our assumption of no estimation error $e$.
Estimation error in the original model was drawn from a normal distribution, and added to the trait value before evaluation of the bias ($A_{ij} = A_j + e_i$).
When estimation error is applied, we sample $J_i$ for each copier $i$ from a normal distribution with expected value zero and variance $\eta^2$. % TODO J_i? e_i?
Even when the estimation error variance is $0.1$, both the fixation probability and fixation time approximations are accurate~(\Cref{fig:hetro_error}). 
We also relaxed our assumption of a uniform bias weight $\alpha$ (i.e., $\alpha_i=\alpha$). We allowed $\alpha$ to vary in the population, drawing $\alpha_j$ for each role-model $j$ from a normal distribution such that $\alpha_j \sim N(0.5,q)$ where $q \in [10^{-7},10^{-1}]$. 
We found again that results of the DM approximation are similar to those from stochastic simulations of the full model~(\Cref{fig:hetro_alpha}).

@@

% TODO I'm not sure this needs to be a "Results" - I think we can write it here as apart of the analysis, which gives Result 4. Also, where is the actual proof of the equation in this result?
\begin{result}[Drift and diffusion terms in a constant environment]
Let $x$ and $x'$ be the frequency of type $\hat{A}$ in a population with $N$ individuals in the current and next generation, and  $\beta$ be the success coefficient of phenotype $A$, $\beta = \beta(A) < \beta(\hat{A}) = 1$.
% TODO can we say beta=beta(A)/beta(hat A) ? then we dont force beta(hat A)=1
Then,
\begin{equation} 
% TODO the use of big O is incorrect... 
E[x'-x] = \mathcal{O} (x(1-x)(1-\beta)) \;, 
\quad
V(x'-x) = \mathcal{O} (x(1-x)\Big(\frac{1}{\alpha N + (1-\alpha)}\Big)) \;.
\end{equation} 
\end{result}

@@

\begin{result}[Fixation time] The expected fixation time (conditioned on fixation) from an initial frequency $x$ is approximately
\begin{equation} \label{eq:kimura_t}
T(x) = J_1(x) + \frac{1-\pi(x)}{\pi(x)}\cdot J_2(x),
\end{equation}
where $N_e=\alpha N + (1-\alpha)$, $S=N_e(1-\beta)$, and
\begin{equation}
\begin{aligned}
J_1(x) &= \frac{-1}{(1-\beta)(e^{-2S}-1)}\int_x^1 \frac{1-e^{2S\xi}-e^{-2S(1-\xi)}+e^{-2S}}{\xi(1-\xi)}d\xi \;, \\
J_2(x) &= \frac{-1}{(1-\beta)(e^{-2S}-1)}\int_0^x \frac{(1-e^{2S\xi})(e^{-2S\xi}-1)}{\xi(1-\xi)}d\xi \;.
\end{aligned}
\end{equation}
\end{result}
Note that these integrals cannot be solved in closed form, {and are estimated} numerically.

@@

\paragraph{Numerical validation.}
To validate the the approximation for the fixation probability in a changing environment (\cref{eq:ch_env}), we compare it to results of simulations that use the DM approximation (\Cref{cor:dirichlet}).
We find that the approximation fits the simulation results well for variable bias weights, $\alpha$, which corresponds to the effective population size (\Cref{fig:ch_env_alpha_beta}A).
